---
title: Computational Approaches to Understanding Surface Heat Flow, the Metamorphic Rock Record, and Subduction Geodynamics
author: Buchanan C. Kerswell
degree: Doctor of Philosophy
discipline: Geosciences
advisor:
  name: Matthew J. Kohn
  title: Ph.D.
  function: Chair, Supervisory Committee
committeeMemberA:
  name: C.J. Northrup
  title: Ph.D.
  function: Member, Supervisory Committee
committeeMemberB:
  name: H.P. Marshall
  title: Ph.D.
  function: Member, Supervisory Committee
externalMember:
  name: Philippe Agard
  title: Ph.D.
  affil: Sorbonne Université
  function: External Member, Supervisory Committee
dean:
  name: Tammi Vacha-Haase
  title: Dean of the Graduate College
date: November 2021
oralexamdate: August 27, 2021
copyright: true
dedication: assets/tex/dedication
acknowledgment: assets/tex/acknowledgment
abstract: assets/tex/abstract
glossary: assets/tex/abbreviations
symbols: assets/tex/symbols
documentclass: report
fontsize: 12pt
colorlinks: true
classoption:
- twoside
biblio-style: assets/bib/authordate1
bibliography: [assets/bib/chpt2ref.bib]
output:
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: pdflatex
    template: assets/tex/bsu_dissertation_template.tex
    toc_depth: 3
    citation_package: natbib
---

```{r settings, echo=FALSE}
# To keep intermediate .tex files like .log
options(tinytex.latexmk.clean = FALSE)
options(tinytex.latexmk.emulation = FALSE)
# Some recommended settings
knitr::opts_chunk$set(
  echo = FALSE,
  message = F,
  fig.pos = 'h',
  out.extra = "", # To force the use of figure enviroment
  fig.align = 'center',
  out.width = '100%'
)
suppressMessages({
  library(dplyr)
  library(purrr)
  library(knitr)
  library(kableExtra)
  library(scales)
  library(pander)
})
```

# Introduction

\markboth{Chapter 1: Introduction}{Chapter 1: Introduction}

> **Keypoints:**
>
> - Proxy datasets are key for inference about geodynamics deep in \glspl{sz}
>
> - Computation leverages large data to infer, build, and test geodynamic models

\cleardoublepage

# Effects of thermo-kinetic boundary conditions on mechanical plate coupling in subduction zones

\markboth{Chapter 2: Coupling Depths}{Chapter 2: Coupling Depths}

```{r load_chpt2, echo=FALSE}
suppressMessages({
  load('assets/r/chpt2_data.RData')
  load('assets/r/chpt2_regressions.RData')
})
```

> **Keypoints:**
>
> - Mechanical coupling responds strongly to \gls{upt}
>
> - Inverting \glsfirst{shf} allows \glsfirst{cd} estimation
>
> - Globally consistent \gls{upt} would support globally uniform \glspl{cd}

## Abstract

Deep mechanical coupling between converging plates is a key feature of \glsfirst{sz} geodynamics. Onset of coupling likely corresponds with metamorphic dehydration reactions, and therefore, is dependent on \glsfirst{pts} fields within \glspl{sz}. Here we consider the effects of changing \glsfirst{tkbc} on coupling using two-dimensional numerical models of oceanic-continental convergent margins. We focus specifically on responses to thermal parameter ($\Phi$) and \glsfirst{upt}. Coupling is implemented numerically by including experimentally-determined (de)hydration reactions of antigorite and olivine. During the experiments, we observe thermal feedbacks regulating (de)hydration self-consistently within the mantle wedge---stabilizing coupling after ca. 5 $Ma$. \glspl{cd} respond strongly to \gls{upt} and weakly to $\Phi$. Regression of our results allows \glsfirst{cd} estimation for modern \gls{sz} segments by inverting \gls{upt} from \glsfirst{shf}. We consider the implications for common \glspl{cd} among \glspl{sz}, which would require consistent \gls{upt}, and thus, globally consistent \gls{shf} in the backarc region.

## Introduction

Deep subduction geodynamics strongly depend on the depth where the subducting plate and overlying mantle transition from mechanically decoupled (moving differentially with respect to each other) to mechanically coupled [moving with the same local velocity, @Furukawa1993; @Peacock1994; @Wada2008]. Traction forces drive mantle wedge circulation at the (de)coupling transition, defining a rapid increase in temperature along the top of the subducting plate [@Peacock1996]. Many observations from numerical experiments and \gls{shf} infer \glspl{cd} occurring globally at 70-80 $km$ in modern \glspl{sz}, essentially independent of other \gls{tkbc} including oceanic-plate age ($t_{OP}$), convergence velocity ($\vec{v}_{conv}$), and subduction geometry [@Furukawa1993; @Wada2008; @Wada2009]. It is significant and curious why modern subduction zones appear to achieve similar depths of coupling despite their different physical characteristics.

Notwithstanding, many numerical geodynamic models use \glspl{cd} of 70-80 $km$ as a boundary condition [e.g., @Abers2017; @Currie2004; @Syracuse2010; @VanKeken2011; @VanKeken2018; @Wada2012; @Gao2014; @Wilson2014], although not exclusively [e.g. 40-56 $km$, @England2010; @Peacock1996]. Similar \glspl{cd} among \glspl{sz} is an attractive hypothesis for at least two reasons: 1) it helps explain the relatively narrow range of sub-arc slab depths [@England2004; @Syracuse2006], as mechanical coupling is expected to be closely associated with the onset of flux melting, and 2) since mechanical coupling is required to detach and recover rocks from the subducting plate [@Agard2016], a common depth of coupling may also help explain why the maximum pressures recorded by subducted oceanic material worldwide is ca. 2.3-2.5 $GPa$ [roughly 80 $km$, @Agard2009].

The location and extent of mechanical coupling along the plate interface is implicated in myriad geodynamic phenomena [seismicity, metamorphism, volatile fluxes into the mantle wedge, volcanism, and plate motions, e.g., @Cizkova2013; @Gonzalez2016; @Peacock1990; @Peacock1991; @Peacock1993; @Peacock1996; @Peacock1999a; @Hacker2003; @VanKeken2011; @Grove2012; @Gao2017]. Consequently, the mechanics of coupling have been extensively studied and discussed. Coupling fundamentally depends on the strength ($\eta$; viscosity) of materials above, within, and below the plate interface. In general, high water fluxes due to compaction and dehydration of clays and other hydrous minerals in the shallow forearc mantle wedge, coupled with increases in \gls{pt}, form layers of low viscosity sheet silicates---especially talc and serpentine---that inhibit transmission of shear stress from the slab to the mantle wedge [@Peacock1999a]. The lack of traction along the interface combined with cooling from the subducting plate surface ensures the shallow mantle wedge remains cold and rigid. Experimentally determined flow laws [e.g., @Agard2016], petrologic observations [e.g., @Agard2016], and geophysical observations [e.g., @Gao2014; @Peacock1999a] all support the plausibility of this conceptual model of subduction interface behaviour.

This chapter focuses on two fundamental questions: 1) how does mechanical \gls{cd} respond to \gls{tkbc}, and 2) how stable is mechanical \gls{cd} through time? We use two-dimensional numerical geodynamic models of subduction to investigate potential correlations between \gls{cd}, \gls{upt} (inverted from backarc heat flow), and the $\Phi$. @Wada2009 previously investigated steady-state slab-mantle \glspl{cd} by modelling 17 active subduction zones. Among other parameters, their models specify convergence rate, subduction geometry, thermal structure of incoming and overriding plate, and degree of coupling along the subduction interface. Their experiments control for interface rheology and discriminate the best-fit depth based on observed fore-arc heat flow. In our models, we similarly specify \gls{tkbc} to simulate the range of modern \gls{sz} systems. However, subduction dip angle and, most importantly, interface rheology are regulated self-consistently by evolving \gls{pts} fields in a deforming mantle wedge. \gls{cd} in each of our models is not a fully determined feature, therefore, but rather a spontaneous model outcome within the range of boundary conditions. As in other previous studies [e.g., @Ruh2015], we include the rheological effect of the dehydration reaction $antigorite \allowbreak \Leftrightarrow olivine + orthopyroxene + H_{2}O$, which drives mechanical coupling by an abrupt viscosity increase with antigorite loss. The position of this reaction along the subduction interface determines the \gls{cd}.

In this study we simulate subduction and observe mechanical plate coupling for 64 convergent margins with variable \gls{upt} and $\Phi$. We quantify \gls{cd} responses to $\Phi$ and \gls{upt} using multi-variate linear regression. We then visualize thermal feedbacks within the system in terms of mantle temperature, viscosity, and velocity fields. Lastly, we discuss how feedbacks stabilize \glspl{cd} through millions of years of active subduction.

\cleardoublepage

# A comparison of heat flow interpolations near subduction zones

\markboth{Chapter 3: Heat Flow Interpolations}{Chapter 2: Heat Flow Interpolations}

> **Keypoints:**
>
> - Inconsistent spatial patterns characterize heat flow near subduction zones
>
> - Heat flow investigations favour 2D interpolations over 1D transects
>
> - Scaling datasets and new interpolation schema will advance \gls{sz} research

## Abstract

Heat fluxing through the Earth's surface provides indirect observations of \glsfirst{pts} fields deep in \glspl{sz}. Global heat flow databases, therefore, are invaluable for generating and testing belief about \gls{sz} geodynamics. Here we argue that investigating \glsfirst{shf} in two-dimensions by interpolation, rather than in one-dimension by projection, forms better interpretations about spatial continuity of deep processes. We directly compare interpolations based on the First (spatial continuity) and Third (similarity) Laws of Geography applied to the most updated global heat flow database. We observe inconsistent spatial patterns and of \gls{shf} in magnitude and variance near subduction zones, regardless of interpolation method. The implications include discontinuous \gls{pts} fields at depth, countering hypotheses of commonly thin upper plate lithospheres and mechanical \glspl{cd} among subduction zones. Strategic scaling of \gls{shf} datasets will improve interpolation precision and confidence---leading to better tools for distinguishing differences within and among \glspl{sz}. We propose new data acquisition and composite interpolation schema as avenues for future \gls{sz} research.

\cleardoublepage

::: {#refs}
:::

\markboth{References}{References}

# (APPENDIX) Appendix {-}
